---
title: "Week 1"
author: "Mirza S. Khan"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
```

# Setup Reticulate

```{r}
library(reticulate)

# Which Python is it inclined to work with?
#py_config()

# Specify which flavor of Python
use_python("/Users/asnasiddiqui/miniconda3/bin/python", required = T)
```


# Load the data

```{r}
data <- vroom::vroom("data/athlete_events.csv")
setDT(data)
```

# Q1

**How old were the youngest male and female participants of the 1992 Olympics?**

```{r}
data[Year == 1992, .(min_age = min(Age, na.rm = T)), by = Sex]

#    Sex min_age
# 1:   M      11
# 2:   F      12
```

# Q2

**What was the percentage of male basketball players among all the male participants of the 2012 Olympics? Round the answer to the first decimal.**

```{r}
# Number of male bball players in 2012
data[Year == 2012 & Sex == "M" & Sport == "Basketball", uniqueN(Name)] #144

# Total number of male participants in 2012 olympics
data[Year == 2012 & Sex == "M", uniqueN(Name)] #5858

round((144 / 5858)*100, 1) #2.5
```

**Answer:** 2.5%

# Q3

**What are the mean and standard deviation of height for female tennis players who participated in the 2000 Olympics? Round the answer to the first decimal.**

```{r}
round( data[Sex == "F" & Year == 2000 & Sport == "Tennis", .(mean = mean(Height, na.rm = T), sd = sd(Height, na.rm = T))] , 1)
# mean: 171.8
# sd: 6.5
```

# Q4

**Find a sportsman who participated in the 2006 Olympics, with the highest weight among other participants of the same Olympics. What sport did he or she do?**

```{r}
data[Year == 2006,][order(Weight, decreasing = T)] %>% 
  dplyr::slice(1) %>% 
  select(Sport) # Skeleton
```

# Q5

**How many times did John Aalberg participate in the Olympics held in different years?**

```{r}
data[Name == "John Aalberg", .(unique(Games))]
#             V1
# 1: 1992 Winter
# 2: 1994 Winter

data[Name == "John Aalberg", uniqueN(Games)]
# 2
```

**Answer:** 2 times

# Q6

**How many gold medals in tennis did sportspeople from the Switzerland team win at the 2008 Olympics? Count every medal from every sportsperson.**

```{r}
data[Team == "Switzerland" & Year == 2008, sum(Medal == "Gold", na.rm = T)] #3
```

```{r}
data %>% 
  filter(Team == "Switzerland" & Year == 2008 & Medal == "Gold") %>% 
  as_tibble()
```

# Q7

**Is it true that Spain won fewer medals than Italy at the 2016 Olympics? Do not consider NaN values in Medal column.**

```{r}
# Number of medals Spain won in '16
data[Team == "Spain" & Year == 2016, sum(!is.na(Medal))]
# 43

# Number of medals Italy won in '16
data[Team == "Italy" & Year == 2016, sum(!is.na(Medal))]
# 70
```

**Answer:** True, Spain won *fewer* medals than Italy inthe 2016 Olympics (43 medals for Spain, 70 total medals for Italy)

# Q8

**What age category did the fewest and the most participants of the 2008 Olympics belong to?**

```{r}
table( data[Year == 2008, cut(Age, breaks = c(15, 25, 35, 45, 55))] )

# (15,25] (25,35] (35,45] (45,55] 
#    7275    5504     623      91 
```

# Q9

**Is it true that there were Summer Olympics held in Atlanta? Is it true that there were Winter Olympics held in Squaw Valley?**

```{r}
data[Season == "Summer" & City == "Atlanta", .(unique(Year))]
# 1996

data[Season == "Winter" & City == "Squaw Valley", .(unique(Year))]
# 1960
```

**Answer:** True, True

# Q10

**What is the absolute difference between the number of unique sports at the 1986 Olympics and 2002 Olympics?**

```{r}
abs( data[Year == 1986, uniqueN(Sport)] - data[Year == 2002, uniqueN(Sport)] )
# 15
```






