---
title: "test"
author: "Mirza Khan"
date: "November 30, 2018"
output: html_document
---

# Converting date and datetime columns in `data.table`

`data.table` does not allow specification of date and datetime columns on read, i.e. with `fread`. Thus, you need to convert these columns to the correct format after the initial read. Below is some code to convert these columns using three different packages: `lubridate`, `fasttime` and `anytime`. Base R also has the `asPOSIXct` function as well.

> tl;dr: use `anytime` and `lubridate` because `fasttime` fails at processing date/datetime before the "epoch" of Jan 1, 1970.
> Guessing the correct datetime format is costly, so when dealing with a large amount of data, might be more efficient to specify the datetime format ahead of time.


```{r}
library(data.table)
library(tidyverse)

example <- tibble(
  date_of_birth = (rep("1954-01-01", 4)),
  vital_type = c('RR', 'P', 'T', 'SBP'),
  vitals = c(14, 80, 35.4, 124), 
  vital_date = c("2018-01-01 17:43:00", "2018-01-01 17:43:00", "2018-01-01 17:43:00", "2018-01-01 17:50:00")
) %>% 
  as.data.table()

example
```


## Select only column names containing the term "date"

When selecting multiple columns, it is often easy to use a regular expression to do so. In this example, I want to modify all of the columns containing the word "date."

There are several packages that allow for pattern matching using regular expressions. I often resort to using the `stringr` package. Base R also has several methods that do the same thing. There is an amazing comparison table of `stringr` and base R functions in the [stringr documentation](https://stringr.tidyverse.org/articles/from-base.html) written by Sara Stoudt that is worth referencing.

```{r}
#using `grep`
grep("date", names(example), value = TRUE)

#using `stringr`
str_subset(names(example), "date")
```

> `data.table` also has some neat pattern matching functionality using `patterns()`, `%like%`, `%chin%`, and `%in%`.

We can now use our regular expression to select the columns of interest as shown below:

```{r}
# Method 1
example[, .SD, .SDcols = names(example) %like% "date"]



# Method 2
cols <- grep("date", names(example), value = TRUE)

#2a
example[, ..cols]

#2b
example[, cols, with = FALSE]
```


## Formatting date-time columns


### lubridate

```{r}
# Simple example - Convert to date time
example[, .(lubridate::as_datetime(vital_date))]
example[, .(lubridate::ymd_hms(vital_date))] # using ymd_hms, should be faster to specify format

example[, .(lubridate::ymd(date_of_birth))]
```

```{r}
cols <- grep("date", names(example), value = TRUE)

example[, (cols) := lapply(.SD, lubridate::as_datetime), .SDcols = cols]
example
```

### anytime

```{r}
example[, .(anytime::anytime(vital_date))]
```

```{r}
cols <- grep("date", names(example), value = TRUE)

example[, (cols) := lapply(.SD, anytime::anytime), .SDcols = cols]
example
```



### fasttime

- Avoid if dealing with dates _before_ Jan 1, 1970

```{r}
# Convert to date time
example[, .(fasttime::fastPOSIXct(vital_date))]
```

```{r}
cols <- grep("date", names(example), value = TRUE)

example[, (cols) := lapply(.SD, fasttime::fastPOSIXct), .SDcols = cols]
example
```

### fasttime failure

> Does `fasttime` not work for dates before the "epoch", i.e. 1970-01-01?

```{r}
fasttime::fastPOSIXct("1969-01-02") # gives NA

as.POSIXct("1969-01-02")
lubridate::as_date("1969-01-02")
anytime::anytime("1969-01-02")
```

