---
title: "Rcpp Notes"
author: "Mirza S. Khan"
date: "7/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basics

C++ data types:

Rcpp includes:

- `*Vector`, e.g. `Rcpp::IntegerVector`, `NumericVector`
- `*Matrix`
- `Rcpp::clone` - make a copy, e.g. `NumericVector v(clone(v_input))`
- Heterogenous datatypes can use `List` or `GenericVector`
- `Rcpp::Named` to assign 'names' to values, e.g. `Rcpp::Named("beta1") = 3`
- `Rcpp::as` (`Rcpp::wrap`) to take data from R to C++ (C++ to R)

* can be `Integer`, `Numeric`, `Logical`, `Character`, `Raw`, any others?

Rcpp Sugar:

- TODO

Use `Rcpp::*Vector`, e.g. `NumericVector`, instead of `std::vector<T>`, where T is the data type, e.g. double. Why? Rcpp vectors do not store copies. 

You can use many functions that act on C++ vectors like their R counterparts, e.g. `max()`.

Avoid Rcpp::*Vector or Rcpp::*Matrix for linear algebra. Use RcppArmadillo instead.

# evalCpp

`evalCpp()` evaluates a single C++ expression. Includes and
dependencies can be declared.

```{r}
Rcpp::evalCpp("2 + 2")

set.seed(42)
Rcpp::evalCpp("Rcpp::rnorm(2)")
```

# cppFunction

`cppFunction()` creates, compiles and links a C++ file, and creates
an R function to access it.

```{r}
Rcpp::cppFunction('
  double add_c(double x, double y) {
    double value = x + y;
    return value;
  }
')

add_c(2,3)
```

# sourceCpp

Write your code in a .cpp file and use the function call `sourceCpp("path/to/file.cpp")`.

Template for the .cpp file:

```
#include <Rcpp.h>
using namespace Rcpp;


// [[Rcpp::export]]
double add_c(double x, double y) {
  double value = x + y;
  return value;
}
```

> **Include `// [[Rcpp::export]]` above each function you want to export/use in R.**

Fun fact: You can use R code blocks in .cpp files; run after compilation.

```
#include <Rcpp.h>
using namespace Rcpp;


// [[Rcpp::export]]
double add_c(double x, double y) {
  double value = x + y;
  return value;
}

// [[Rcpp::export]]
int g(int n) {
	if (n < 2) return(n);
	return(g(n-1) + g(n-2));
}

// [[Rcpp::export]]
double getMax(NumericVector v) {
	return( max( v ) );
}



/*** R
add_c(21, 21)
*/

/*** R
sapply(0:10, g)
*/
```

# Example: Matrix operations

```
#include <Rcpp.h>

// [[Rcpp::export]]
Rcpp::NumericVector colSums(Rcpp::NumericMatrix mat) {
	size_t cols = mat.cols();
	Rcpp::NumericVector res(cols);
	for (size_t i=0; i<cols; i++) {
		res[i] = sum(mat.column(i));
	}
	return(res);
}
```

# Example sum a vector of integers

```{Rcpp}
#include <algorithm>
#include <Rcpp.h>

// [[Rcpp::export]]
int accumulate(Rcpp::IntegerVector v){
  return std::accumulate(v.begin(), v.end(), 0);
}

/*** R
accumulate(1:5)
*/
```



# Create an Rcpp package

1. `Rcpp.package.skeleton()` or `RcppArmadillo.package.skeleton()`
2. RStudio: File -> New Project -> Package menus


# Resources

Resources:
1. http://dirk.eddelbuettel.com/papers/useR2019_rcpp_tutorial.pdf
2. https://csgillespie.github.io/efficientR/rcpp.html
3. http://dirk.eddelbuettel.com/code/rcpp/Rcpp-quickref.pdf
4. https://gallery.rcpp.org/
5. Book: Seamless R and C++ Integration with Rcpp
6. "Unofficial" API documentation - https://github.com/coatless/rcpp-api/blob/master/rcpp-api-docs.Rmd
7. Advanced R: Rcpp chapter - http://adv-r.had.co.nz/Rcpp.html
8. https://youtu.be/xiqYaHa2x4s

