---
title: "Reticulate"
author: "Mirza S. Khan"
date: "7/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE,
                      message = FALSE)
```

## Resources

- [useR 2019 tutorial](https://3mmarand.github.io/useR2019_tutorial/)
  - Features an awesome example of using R and Python [here](https://3mmarand.github.io/useR2019_tutorial/music_ml/Classifiying_music.html)
  
## Setup

```{r}
library(reticulate)

# Which Python is it inclined to work with?
py_config()

# Specify which flavor of Python
use_python("/Users/asnasiddiqui/miniconda3/bin/python", required = T)

# Confirm right Python is selected
Sys.which("python")

py_discover_config()
```

The **4** ways to use Python with `reticulate`:

1. `repl_python()`
2. In Rmd docs, e.g. python chunks
2. Source Python scripts
3. Import Python modules

> Access the things made in Python session from R console/chunk using `py$<thing>`
> The inverse of this is to acces R objects in a Python console/chunk using `r.<thing>`

## REPL

You can use Python in the RStudio console using `repl_python()`. Exit out by pressing the 'Esc' key or by typing `exit` in the console.

```{r}
repl_python()
```

## Python Chunks

```{python}
import numpy as np

x = np.array([1,2,3])
x*2
```

More recently, we are able to carry the variables forward to subsequent chunks. For example, the variable $x$ is available in the next chunk for computation.

```{python}
x*3

x.max()
```

The array we made in python, "x", is also available in our R session using `py$x`.

```{r}
py$x * 2

min(py$x)

# Make available as 'x' in R environment
x <- py$x
```

This awesomeness that is `reticulate` also allows us to use functions and methods from Python in the R environment. For example, use numpy to calculate the max of the array.

```{r}
py$np$max(py$x)
```


## Imports Example

```{r}
library(tidyverse)

boston <- MASS::Boston
```

```{python}
import pandas as pd
import numpy as np
from sklearn.linear_model import LinearRegression

r.boston.shape
r.boston.head()
type(r.boston) # pandas DF
```



